---
title: "R Notebook"
output:
  pdf_document: default
  html_notebook: default
authors: Andreas M. Brandmaier, Rogier Kievit, & Ross Jacobucci
---

In this tutorial, we are going to demonstrate the relcs package. Using relcs, you are able to easily specify random effects latent change score models that allow proportional change (or, self-feedback) to vary between persons. 

First, download the package from GitHub and attach it to your workspace. Here we also load a couple of other libraries that will make our life less miserable.

```{r init, eval=TRUE,echo=TRUE,message=FALSE,warning=FALSE}
require("relcs")

require("ggplot2")
library("ggthemes")
library(tidyr)

set.seed(543231)
```

Let's start with simulating some data from a latent change score model. Here, we assume proportional change (mean=.2) and individual differences in the proportional change (variance=.1). We assume no additional linear slope over time. Residual error variance is set to 1. There is some individual differences at the intercept (mu=5, variance=1).

```{r}
N <- 100
num.obs <- 5

simulated.data <- relcs::simulateDataFromRELCS(N,num.obs,selffeedback.mean = .2, 
                                               selffeedback.variance = .05,has.slope = FALSE,
                                               residualerrorvariance = 1,
                                              interceptmu = 5, interceptvariance = 1)

summary(simulated.data)
```

Let's first create a long version of our simulated data set, which we will use for plotting shortly. The tidyr package is a great helper for this task.
```{r shape,message=FALSE}
simulated.data <- cbind(ID=rep(1:N), simulated.data)
long.simulated.data <- gather(data = simulated.data,X,value,-ID)
long.simulated.data$ID <- as.factor(long.simulated.data$ID)
```

We'll use the amazing ggplot2 library to plot the data

```{r plotstuff,message=FALSE}

subdata <- long.simulated.data
ggplot(data=subdata,aes(x=X,y=value,color=ID,group=ID))+geom_line()+   guides(colour=FALSE)+
  theme_fivethirtyeight()
```

Now, we will fit a RELCS model. Yes, this is really all you need to do:

```{r echo=TRUE, eval=FALSE}
fitted.model <- fitRELCS(data=simulated.data)
```

```{r run_model,echo=FALSE,eval=TRUE,message=FALSE,warning=FALSE,results='hide'}
fitted.model <- fitRELCS(data=simulated.data, iter=200, warmup=40)
```



Yes, it is as simple as that! Let's have a look at the parameter point estimates (here, the means of the posterior distribution):

```{r}
getEstimates(fitted.model) 
```

Now, let's plot our estimates using relcs's in-built plotting functions.

```{r}
 est <- getEstimates(fitted.model)
 plot(est)
```

Now, realistically, your data will have some missing values here and there. Let's simulate a matrix that reflects the pattern of missing and non-missing values and remove some of our observations according to it. We set the probability of missing data to 10% here. 

```{r}
prob.miss <- 0.1 
miss.mat <- ifelse(matrix(data = rbinom(n = prod(dim(simulated.data)),size = 1,prob=prob.miss),nrow=nrow(simulated.data)),NA,1)
missing.data <- miss.mat * simulated.data
```

And again, we simply pass this data object to fitRELCS and it will handle the missing data appropriately:

```{r run_model_missing2, eval=FALSE,echo=TRUE,message=FALSE,warning=FALSE}
require(relcs)
fitted.model <- fitRELCS(data=missing.data, type="stan")
```


```{r run_model_missing, eval=TRUE,echo=FALSE,message=FALSE,warning=FALSE,results='hide'}
require(relcs)
fitted.model <- fitRELCS(data=missing.data, type="stan")
```


Let's again check the estimates using print and plot:

```{r}
 est <- getEstimates(fitted.model)
print(est)
plot(est)
```

