relcs
======

```{r echo=FALSE}
knitr::opts_chunk$set(
  comment = "#>",
  collapse = TRUE
)
```

## What is this?

Random Effects Latent Change Score models.

## Install

Obtain the latest package version from github:

```{r eval=FALSE}
install.packages("devtools")
devtools::install_github("brandmaier/relcs")
```

Load the `relcs` library:

```{r warning=FALSE}
library("relcs")
```

## How To Use

The `relcs` package offers functions to simulate data (either from an OpenMx model or from specialized code) and to fit a model using either STAN or OpenMx.

### Simulation

Simulate 100 cases from a RELCS model:

```{r message=FALSE}
library(relcs)
library(tidyverse)
library(ggplot2)

simulated_data <- simulateDataFromRELCS(N = 100, 
                                        num.obs = 5, 
                                        residualerrorvariance = .3,
                                        selffeedback.mean = .5,
                                        selffeedback.variance = .01,
                                        interceptmu = 0,
                                        interceptvariance = 1,
                                        has.slope=FALSE)
```

Plot the first 20 simulated trajectories:

```{r}
simulated_data %>% 
  mutate(id=1:nrow(simulated_data)) %>% 
  filter(id < 20) %>%
  pivot_longer(-id) %>%
  ggplot(aes(x=name,y=value,group=id,color=factor(id)))+
  geom_line()+
  theme_minimal()+
  xlab("Time")+ylab("Value")+
  ggx::gg_("hide legend")
```

### model fit

```{r warning=FALSE, echo=TRUE}
fit <- fitRELCS(data = simulated_data, type="stan")

fit$fit
```

### visually inspect model posteriors

```{r}
hist(rstan::extract(fit$fit)$residual_var,main = "Residual Variance")
```
### Obtain individual beta estimates

Re-fit the model with option `beta.as.parameter` which makes the person-specific parameters
additional parameters, one for each person.

```{r warning=FALSE}
fit_with_beta <- fitRELCS(data = simulated_data, type="stan",beta.as.parameter = TRUE)
```

Get the beta values from the model and obtain summary.

```{r warning=FALSE}
betas <- get_beta_estimates(fit_with_beta)
summary(betas)
hist(betas)
```

### inspect STAN code

```{r}
cat(fit$code)
```